The Array of Long-Baseline Antennas for Taking Radio Observations from the Sub-Antarctic

### 2020-06-09 albatros_paper.tex
HCC tweaked schematic figures, added more 2-element text

\begin{figure}
  \begin{center}
    \includegraphics[width=0.7\linewidth]{Figures/albatros_2elem/albatros_2elem.pdf}
    \caption{The two-element, directly correlated
      \albatros\ pathfinder installed at the \prizm\ site.  Two
      dual-polarization antennas are separated by roughly 110~m on an
      east--west baseline. Coaxial cables connect the antennas to a
      shipping container that houses the readout electronics and
      serves as the ``command module.'' \attention{Add FEE
        and backend sub-pics here? (HCC)}}
    \label{Fig:albatros2}
  \end{center}
\end{figure}

\begin{figure}
  \begin{center} \includegraphics[width=1.0\linewidth]{Figures/albatros_2elem_schematic/albatros_2elem_schematic.pdf}
    \caption{Two-element \albatros\ pathfinder block diagram.  Signals
      from two dual-polarzation LWA antennas are amplified by
      front-end active baluns~\citep{2012PASP..124.1090H}.  Antennas
      are connected via 100m coax cables to the back-end readout
      electronics, which are housed in a Faraday cage denoted by the
      dashed box.  Each of the four antenna outputs is passed to a
      second-stage electronics chain consisting of filters and further
      amplfication.  The signals are digitized at 250~Msamp/s by a
      SNAP board, which includes an on-board FPGA that computes auto
      and cross spectra from and between the four inputs.  A Raspberry
      Pi controls the SNAP board and saves the data.}
    \label{Fig:albatros2_schem}
  \end{center}
\end{figure}

The first exploratory \albatros\ measurements were conducted with a
two-element pathfinder that employed direct correlation (without
autonomous operation).  \autoref{Fig:albatros2} illustrates this
pathfinder, which was installed at the \prizm\ site (\ang{46;53;13}S,
\ang{37;49;10.7}E) in April 2018.  The system block diagram is shown
in \autoref{Fig:albatros2_schem}, and each of the subsystems is
described in detail below.

\subsection{Antenna}	
We employ two dual-polarization Long Wavelength Array (LWA) dipole
antennas~\citep{Memo28}.  The LWA antennas have a long development
history, are well characterized, and are simple to install and
physically robust.  The antennas form an east--west baseline with a
separation of \SI{110}{m}, and the polarizations are aligned with the
cardinal directions.  Welded wire mesh screens, roughly 3~m on a side,
are installed on the ground below the antennas.
% do we want/need to say anything about the omnidirectional primary beam?

\subsection{Front-end active balun}
\attention{Tankiso, Eamon, Jeff?} \\
\label{section:FEE}

The LWA active-balun front-end electronics (FEE) circuit uses a MiniCircuits GALI-74+ MMIC 
to amplify each dipole leg against ground, presenting each leg with a 50$\ohm$ impedance 
and providing a nominal gain of 25dB. The two GALI-74+ outputs are differenced using a 
passive {180\degree} hybrid coupler. The coupler output is filtered by a $\sim$150MHz lowpass, 
and receives an additional 12dB of gain from a MiniCircuits GALI-6+ MMIC. This last amplifier 
drives the output signal onto a 100m 50$\ohm$ coaxial cable. Thus, neglecting mismatch loss 
between the dipoles (electrically small in the frequency range of interest) and the 100{\ohm} 
input impedance, as well as the hybrid insertion loss ($<$1dB), the front-end electronics 
provide a gain of $\sim$37~dB~\citep{2012PASP..124.1090H}.

Each FEE circuit is powered by 16V, fed on the coax through a bias tee.

@@

% It seems to me that this is too much detail to go into in describing an already published design.
\iffalse

All the front end components were incorporated for in a double sided
printed circuit board (PCB) as shown in \autoref{Fig:Balun} and the
block diagram is shown in \autoref{Fig:Balun Schematic}. The
Monolithic Microwave Integrated Circuits (MMICs) is the used design
for the PCB. One side of the PCB is populated with components and the
other side is a solid copper ground plane aperiodically stitched to
the grounded copper on the side populated with components. The
receiver system is made up of the active balun, filter and the gain
stage that connects to the \SI{100}{\metre} coaxial cable which is
connected to the back end \cite{2012PASP..124.1090H}.

The input impedance (Z\textsubscript{o}) of \SI{50}{\ohm} is
introduced to the dipole by the active balun. The signal is then fed
through an amplifier which amplifies it by \SI{+24}{\decibel} of
gain. The balanced signal is then converted to unbalanced through a
180 \degree hybrid coupler. The band pass filter (BPF) receives the
single ended signal in order for it to reject all the frequencies
which are not within the range of interest. The signal gets fed to a
second amplifier which again amplies it by \SI{+24}{\decibel} of gain
and the output impedance of the FEE is matched to a \SI{50}{\ohm}
coaxial cable. The bias tee is responsible for providing power to the
FEE and extracts the RF signal by the use of the coaxial cable. This
unit has an overall gain of $\approx$ \SI{35}{\decibel} and an overall
noise figure of $\approx$ \SI{2.7}{\decibel} to $\approx$
\SI{2.9}{\decibel} \cite{Memo35}.

\fi

\subsection{Back-end electronics}
\attention{Nivek, HCC?} \\

The back-end readout electronics are housed in a Faraday cage
located 100m away from the antennas to mitigate possible
self-generated RFI.  Each of the four antenna
signals is passed to a second-stage electronics chain consisting of 
% EE bias tee is already described; it isn't really part of the signal chain since it doesn't do much of anything
%a bias tee (Minicircuits ZFBT-4R2GW-FT+), 
an
amplifier (Minicircuits
ZX60-V63+), and a pair of high- and low-pass filters (Minicircuits
ZFHP-1R2+ and SLP-90+) that together  bandlimit the signal to
1.2--\SI{81}{MHz}.  The amplifier has a nominal gain of 20dB, and the HPF and LPF contribute nominal insertion losses of 0.2dB and 0.14dB respectively.

A Xilinx Kintex\footnote{http://www.xilinx.com/products/silicon-devices/fpga/kintex-7.html} based Smart Network ADC Processor (SNAP) board (\attention{citation for SNAP board?}), 
clocked by a Valon 5007 frequency synthesizer module,
digitizes the RF signals at 250~Msamp/s. It calculates a full
cross-correlation of the four inputs, producing four auto- and six
cross-spectra as outputs, over 2048 channels spanning the frequency range 
\SI{0}{MHz} - \SI{125}{MHz}. 
(\attention{details about SNAP board frequency transform?})
The SNAP board sends its output data via ethernet to a Raspberry Pi (\attention{version? citation?}), which saves baseband data to a set of hard drives, while storing the less voluminous correlation data on its SD card.

@@

A bank of four parallel \SI{12}{V}, \SI{200}{\ampere\hour} batteries serves to power the two-element pathfinder system. A Honda EU30is generator is run periodically to charge the battery bank, which can run the system for approximately four days on a full charge.  DC/DC converters power the SNAP board, the FEE, the amplifiers and the clock.
% Cuts (redundant or too much detail)
% and the fuel is kept in one of the shipping containers at the \prizm\ site
% The power from the battery bank is connected to the Faraday cage where the power regulation is performed.
\subsection{Software and data acquisition}
\attention{Nivek?} \\
Text goes here.

@@

% single element block diagram goes here
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{Figures/albatros_single_schematic/albatros_single_schematic.pdf}
    \caption{Single-antenna autonomous station block diagram. 
      Data and RF connections are outlined in black, and power
      connections in red.  
      A dual-polarization LWA
      antenna, equipped with an active balun (described above in \ref{section:FEE}), connects via 50m
      coaxial cables to the back-end readout electronics, 
      housed in a Faraday cage denoted by the dashed box. 
      The antenna signals are passed
      to a second-stage electronics chain consisting of filters and further amplfication.  The signals are
      digitized at 250~Msamp/s by a SNAP board, which includes an on-board FPGA that computes spectra from both antennas.  
      A Raspberry Pi controls the SNAP board, receives the spectrum of interest from the SNAP, and saves it as baseband data.}
    \label{Fig:Signal Chain}
  \end{center}
\end{figure}
\attention{Nivek?} \\

>>>>

\begin{figure}
  \begin{center}
    \includegraphics[width=0.7\linewidth]{Figures/albatros_2elem/albatros_2elem.pdf}
    \caption{The two-element, directly correlated
      \albatros\ pathfinder installed at the \prizm\ site.  Two
      dual-polarization antennas are separated by roughly 110~m on an
      east--west baseline. Coaxial cables connect the antennas to a
      shipping container that houses the readout electronics and
      serves as the ``command module.''}
    \label{Fig:albatros2}
  \end{center}
\end{figure}

\begin{figure}
  \begin{center} \includegraphics[width=1.0\linewidth]{Figures/albatros_2elem_schematic/albatros_2elem_schematic.pdf}
    \caption{Two-element \albatros\ pathfinder block diagram.  Signals
      from two dual-polarzation LWA antennas are amplified by
      front-end active baluns~\citep{2012PASP..124.1090H}.  The
      antennas are connected via 100-m coaxial cables to the back-end
      readout electronics, which are housed in a Faraday cage denoted
      by the dashed box.  Each of the four antenna outputs is passed
      to a second-stage electronics chain consisting of filters and
      further amplfication.  The signals are digitized at 250~Msamp/s
      by a SNAP board, which includes an on-board FPGA that computes
      auto- and cross-spectra from and between the four inputs.  A
      Raspberry Pi controls the SNAP board and saves the data.}
    \label{Fig:albatros2_schem}
  \end{center}
\end{figure}

The first exploratory \albatros\ measurements were conducted with a
two-element pathfinder that used direct correlation (without
autonomous operation).  \autoref{Fig:albatros2} illustrates this
pathfinder, which was installed at the \prizm\ site (\ang{46;53;13}S,
\ang{37;49;10.7}E) in April 2018.  The system block diagram is shown
in \autoref{Fig:albatros2_schem}, and each of the subsystems is
described in detail below.

\subsection{Antenna}

We employ two dual-polarization Long Wavelength Array (LWA) dipole
antennas~\citep{Memo28}.  The LWA antennas have a long development
history, are well characterized, and are simple to install and
physically robust.  The antennas form an east--west baseline with a
separation of \SI{110}{m}, and the polarizations are aligned with the
cardinal directions.  Welded wire mesh screens, roughly 3~m on a side,
are installed on the ground below the antennas.
% do we want/need to say anything about the omnidirectional primary beam?

\subsection{Front-end active balun}\label{section:FEE}

The LWA active-balun front-end electronics (FEE) circuit uses a
MiniCircuits GALI-74+ MMIC to amplify each dipole leg against ground,
presenting each leg with a 50$\ohm$ impedance and providing a nominal
gain of 25~dB. The two GALI-74+ outputs are differenced using a
passive {180\degree} hybrid coupler. The coupler output is filtered by
a $\sim$150~MHz lowpass and receives an additional 12~dB of gain from
a MiniCircuits GALI-6+ MMIC. This last amplifier drives the output
signal onto a 100-m 50$\ohm$ coaxial cable. Thus, neglecting mismatch
loss between the dipoles (electrically small in the frequency range of
interest) and the 100{\ohm} input impedance, as well as the hybrid
insertion loss ($<$1dB), the front-end electronics provide a gain of
$\sim$37~dB~\citep{2012PASP..124.1090H}.  Each FEE circuit is powered
by 16~V, which is fed on the coaxial cable through a bias tee.

@@

%% It seems to me that this is too much detail to go into in describing an already published design.
% All the front end components were incorporated for in a double sided
% printed circuit board (PCB) as shown in \autoref{Fig:Balun} and the
% block diagram is shown in \autoref{Fig:Balun Schematic}. The
% Monolithic Microwave Integrated Circuits (MMICs) is the used design
% for the PCB. One side of the PCB is populated with components and
% the other side is a solid copper ground plane aperiodically stitched
% to the grounded copper on the side populated with components. The
% receiver system is made up of the active balun, filter and the gain
% stage that connects to the \SI{100}{\metre} coaxial cable which is
% connected to the back end \cite{2012PASP..124.1090H}.

% The input impedance (Z\textsubscript{o}) of \SI{50}{\ohm} is
% introduced to the dipole by the active balun. The signal is then fed
% through an amplifier which amplifies it by \SI{+24}{\decibel} of
% gain. The balanced signal is then converted to unbalanced through a
% 180 \degree hybrid coupler. The band pass filter (BPF) receives the
% single ended signal in order for it to reject all the frequencies
% which are not within the range of interest. The signal gets fed to a
% second amplifier which again amplies it by \SI{+24}{\decibel} of gain
% and the output impedance of the FEE is matched to a \SI{50}{\ohm}
% coaxial cable. The bias tee is responsible for providing power to the
% FEE and extracts the RF signal by the use of the coaxial cable. This
% unit has an overall gain of $\approx$ \SI{35}{\decibel} and an overall
% noise figure of $\approx$ \SI{2.7}{\decibel} to $\approx$
% \SI{2.9}{\decibel} \cite{Memo35}.

\subsection{Back-end electronics}

The back-end readout electronics are housed in a Faraday cage located
100~m away from the antennas to mitigate possible self-generated RFI.
Each of the four antenna signals is passed to a second-stage
electronics chain consisting of an amplifier (Minicircuits ZX60-V63+),
and a pair of high- and low-pass filters (Minicircuits ZFHP-1R2+ and
SLP-90+) that together band-limit the signal to 1.2--\SI{81}{MHz}.
The amplifier has a nominal gain of 20~dB, and the high- and low-pass
filters contribute nominal insertion losses of 0.2~dB and 0.14~dB,
respectively.

A Smart Network ADC Processor~\citep[SNAP;][]{2016JAI.....541001H}
board digitizes the RF signals at 250~Msamp/s and uses a Xilinx
Kintex~7\footnote{http://www.xilinx.com/products/silicon-devices/fpga/kintex-7.html}
FPGA to calculate full cross-correlations of the four inputs,
producing four auto- and six cross-spectra as outputs, over 2048
channels spanning the frequency range 0--125~MHz.  \attention{details
  about channelization here} A Valon 5007 frequency synthesizer module
provides the clock signal for the SNAP board.  A Raspberry Pi~3B+
(RPi) single board computer controls the SNAP board and receives the
spectra via GPIO connections, and the spectra are saved to an on-board
SD card.  An Adafruit Ultimate GPS
module\footnote{\url{https://www.adafruit.com/product/746}}, connected
to an active external GPS antenna, provides absolute timing for the
RPi.

@@

A bank of four 12-V, 200-Ah AGM batteries, connected in parallel,
powers the two-element pathfinder system.  The batteries are manually
charged with a Honda EU30is generator that is housed on site.  The
total power draw of the two-element pathfinder is $\sim45$~W, and when
fully charged, the battery bank can power the system for about a week.
The raw battery voltage is passed to several DC/DC converters that
supply regulated voltages to the SNAP board, FEE, amplifiers and the
clock.

\subsection{Software and data acquisition}
\attention{Nivek?} \\
Text goes here.

@@

% single element block diagram goes here
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{Figures/albatros_single_schematic/albatros_single_schematic.pdf}
    \caption{Single-antenna autonomous station block diagram. 
      A dual-polarization LWA
      antenna, equipped with an active balun (described above in \ref{section:FEE}), connects via 50m
      coaxial cables to the back-end readout electronics, 
      housed in a Faraday cage denoted by the dashed box. 
      The antenna signals are passed
      to a second-stage electronics chain consisting of filters and further amplfication.  The signals are
      digitized at 250~Msamp/s by a SNAP board, which includes an on-board FPGA that computes spectra from both antennas.  
      A Raspberry Pi controls the SNAP board, receives the spectrum of interest from the SNAP, and saves it as baseband data.}
    \label{Fig:Signal Chain}
  \end{center}
\end{figure}
\attention{Nivek?} \\
